Luminance constancy under fixed geometry

### 2018-01-29 Draft2/draft2.tex
more edits
lost of edits

% WHAT STORY DO WE WANT TO TELL ABOUT THE RESULTS
%  1) Case 0 -- work out effect of sources of internal noise - isomerization and AMA -- without any external "noise", where external "noise" is the effect of variability in object extrinsic scene factors (illumination, surrounding objects). 
%                      Also establish the rendering noise is einsy relative to whatever else we're doing.
%  2) Move to Case 1 -- what happens when we introduce color variability into the target.  This may not be of interest to the psychophysics, but none-the-less seems interesting to know as one ponders lightness constancy.
%  3) On to Case 2 -- here the question is, do we keep varying target color, or do we fix target color and vary illumination?  In any case, this isn't too hard a case if the other objects remain fixed, because then contrast coding basically normalizes out a lot of the variation, and it's only the spectral variation that might screw things up.  But, that's why we run case 3.
%  4) Do we run a case where the relative spectrum of the illuminant stays fixed and we only vary intensity?
%  4) On to Case 3 --
\section{Results} \label{Results}
\begin{itemize}  
\item Show the minimal model (to describe the sources and magnitude of error). This can go in supplement.
\begin{itemize}  
\item rendering error
\item isomerization error
\item ama error
\end{itemize}
\item Case 1. Only the target object color changes. AMA on isomerization
\item Case 2. Background color changes too, AMA on isomerization. (This is not giving good results. The filters are giving weights to the background which they should not. Maybe more samples would help.)
\item Case 3. Background fixed, illuminant changes, AMA on isomerization fails, but the contrast version works
\item Case 4. Change all three and show that both ismerization fails, constrast does better, but not perfectly better

\end{itemize}

%% Figure Results
\begin{figure}
\centering
    \begin{subfigure}[b]{0.33 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/resultsFigure/renderingNoise.pdf}
        \caption{Rendering noise}
        \label{fig:renderingNoise}
    \end{subfigure}
        \begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/resultsFigure/renderingAndIsomerization.pdf}
        \caption{Rendering and isomerization noise}
        \label{fig:renderingAndIsomerization}
    \end{subfigure}
        \begin{subfigure}[b]{0.33 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/resultsFigure/renderingIsomerizationTargetColor.pdf}
        \caption{Target color variation}
        \label{fig:renderingIsomerizationTargetColor}
    \end{subfigure}    
    \caption{Different sources of noise}
\label{fig:sourcesOfNoise}
\end{figure}

@@

%% Figure 9
\begin{figure}
\centering
    \begin{subfigure}[b]{0.4 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_a.pdf}
        \caption{AMA First 2 RFs}
        \label{fig:case9AMA}
    \end{subfigure}
        \begin{subfigure}[b]{0.27 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_b.pdf}
        \caption{AMA RF Response}
        \label{fig:case9FiltersResponse}
    \end{subfigure}
        \begin{subfigure}[b]{0.27 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_c.pdf}
        \caption{Luminance Estimates}
        \label{fig:case9Results}
    \end{subfigure}    
    \caption{{\bf Receptive fields and luminance estimates. Case 1:} (a) First two receptive fields produced by the AMA. Along the row we show the filters corresponding to the L cones, M cones and S cones (left to right). (b) The response of the stimuli to the first two AMA filters. The responses are grouped with the label of the stimuli. As is clear, the response distributions at each luminance level separate out well in this case, resulting in good estimates of the target object luminance. (c) The estimates of the target object standard luminance. The naive method produces the mean value of the target luminance used in the analysis. The shaded region corresponds to 1 sigma standard deviation.}
\label{fig:case9AllResults}
\end{figure}

%% Figure Isomerization Fails
\begin{figure}
\centering
    \begin{subfigure}[b]{0.3	 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/resultsFigure/isomerizationFails.pdf}
        \caption{AMA on isomerization fails in case of variation in illuminant}
        \label{fig:isomerizationFails}
    \end{subfigure}
    ~
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/resultsFigure/constrastWorks.pdf}
        \caption{Converting signal to contrast solves the problem}
        \label{fig:constrastWorks}
    \end{subfigure}
    \caption{Need for contrast conversion.}
\label{fig:importanceOfConstrast}
\end{figure}

%% Figure 10
\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{../Figures/Figure10/Figure10.pdf}
\caption{{\bf Performance with number of RFs used for estimation:} Most of the information is captured by the first two filters. We have used the first 8 RFs for estimating the luminance.}
\label{fig:RMSEvsNFilters}
\end{figure}
% GET RID OF AMA FROM THE CAPTION

% Figure 11
\begin{figure}
\centering
    \begin{subfigure}[b]{0.27 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure11/Figure11_a.pdf}
        \caption{AMA First 2 RFs}
        \label{fig:case10AMA}
    \end{subfigure}
        \begin{subfigure}[b]{0.20 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure11/Figure11_b.pdf}
        \caption{AMA RF Response}
        \label{fig:case10FiltersResponse}
    \end{subfigure}    
        \begin{subfigure}[b]{0.2 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure11/Figure11_c.pdf}
        \caption{Luminance Estimates}
        \label{fig:case10Results}
    \end{subfigure}
    \caption{{\bf Receptive fields and luminance estimates. Case 2:} Same as Fig. \ref{fig:case9AllResults}.}
    \label{fig:case10AllResults}
\end{figure}

% Figure 12
\begin{figure}
\centering
    \begin{subfigure}[b]{0.27 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_a.pdf}
        \caption{AMA First 2 RFs}
        \label{fig:case12AMA}
    \end{subfigure}
        \begin{subfigure}[b]{0.20 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_b.pdf}
        \caption{AMA RF Response}
        \label{fig:case12FiltersResponse}
    \end{subfigure}    
        \begin{subfigure}[b]{0.2 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_c.pdf}
        \caption{Luminance Estimates}
        \label{fig:case12Results}
    \end{subfigure}
    \caption{{\bf Receptive fields and luminance estimates. Case 3:}  Same as Fig. \ref{fig:case9AllResults}.}
    \label{fig:case12AllResults}
\end{figure}

@@

\subsection{The RFs of complex case generalize on simpler cases}
% Figure 13
\begin{figure}
\centering
\begin{subfigure}{0.3 \textwidth}
	\includegraphics[width=\textwidth]{../Figures/Figure13/Figure13_a.pdf}
	\caption{Case 1}
	\label{fig:case1Bar}
    \end{subfigure}
    ~ 
    \begin{subfigure}{0.3 \textwidth}   
	\includegraphics[width=\textwidth]{../Figures/Figure13/Figure13_b.pdf}
	\caption{Case 2}
	\label{fig:case2Bar}
    \end{subfigure}
    ~ 
        \begin{subfigure}{0.3 \textwidth}
	\includegraphics[width=\textwidth]{../Figures/Figure13/Figure13_c.pdf}
	\caption{Case 3}
	\label{fig:case3Bar}
    \end{subfigure}
\caption{{\bf Model and Filter Performance:} Comparison of optimal AMA RFs from one case on the images of other cases. (a) RMSE of luminance estimates on images from the training set of case 1. Each bar represents RMSE of the filter obtained using the optimal RFs corresponding to the model. (b) (c)}
 \label{fig:barGraphs}
\end{figure}
Fig.~\ref{fig:SVDBAR} and Fig~\ref{fig:AMABAR} show the performance of the receptive fields learnt using the stimuli of one case on the stimuli of other cases. The RFs of case 3, (the most complex case) perform well on stimuli of case 1 and case 2.  Similarly, the RFs of case 2 performs well on the stimuli of case 1. Thus, the receptive fields leant on the complex case generalize for luminance estimation of stimuli for simpler cases.

>>>>

% WHAT STORY DO WE WANT TO TELL ABOUT THE RESULTS
%  1) Case 0 -- work out effect of sources of internal noise - isomerization and AMA -- without any external "noise", where external "noise" is the effect of variability in object extrinsic scene factors (illumination, surrounding objects). 
%                      Also establish the rendering noise is einsy relative to whatever else we're doing.
%  2) Move to Case 1 -- what happens when we introduce color variability into the target.  This may not be of interest to the psychophysics, but none-the-less seems interesting to know as one ponders lightness constancy.
%  3) On to Case 2 -- here the question is, do we keep varying target color, or do we fix target color and vary illumination?  In any case, this isn't too hard a case if the other objects remain fixed, because then contrast coding basically normalizes out a lot of the variation, and it's only the spectral variation that might screw things up.  But, that's why we run case 3.
%  4) Do we run a case where the relative spectrum of the illuminant stays fixed and we only vary intensity?
%  4) On to Case 3 --
\section{Results} \label{Results}
\begin{itemize}
\item Show the minimal model (to describe the sources and magnitude of error). This can go in supplement.
\begin{itemize}  
\item rendering error
\item isomerization error
\item ama error
\end{itemize}
\item Case 1. Only the target object color changes. AMA on isomerization
\item Case 2. Background color changes too, AMA on isomerization. (This is not giving good results. The filters are giving weights to the background which they should not. Maybe more samples would help.)
\item Case 3. Background fixed, illuminant changes, AMA on isomerization fails, but the contrast version works
\item Case 4. Change all three and show that both ismerization fails, constrast does better, but not perfectly better
\end{itemize}

\subsection{Case 1: Target luminance can be estimated directly from the photoreceptor response}
We used AMA to learn a set of linear receptive fields that are optimal for estimating target object luminance using stimuli in the training set of case 1. 
Fig.~\ref{fig:case1RFResponse} shows the projection of the photoreceptor response along the first two AMA receptive fields. 
Clearly, there exists a subspace along which the photoreceptor response separates according to the luminance level assigned to the target object. 
Thus, the target object luminance can be obtained by projecting photoreceptor responses along these AMA receptive fields. 
Fig.~\ref{fig:case9Results} shows the estimated luminance v/s the actual luminance assigned to the target object for the test images in case 1. 
We show the estimates obtained by AMA and two null models: a naive method that assigns the average target object luminance across the training set to all images, and a linear model based on the photoreceptors corresponding to the center pixel on the target object (see Methods). 
As expected, in the absence of any external source of variation, the photoreceptor response is proportional to the luminance of the target object, and hence can be easily estimated by both AMA and linear regression.

Each cloud in Fig.~\ref{fig:case1RFResponse} corresponds to the photoreceptor response to the images with target luminance level at a fixed value. 
The variation within the cloud, between images that have the target object at same luminance values, results due to the difference in the relative shape of the target object reflectance spectra (difference in hue and saturation). 
The resulting variation dominates any variation coming from the noise in photoreceptor response or the noisy rendering of the images by the software.
Fig.~\ref{fig:case0RFResponse} shows the projection of photoreceptor response corresponding to 1000 images where, in addition to the illumination and background spectra, we also fixed the relative shape of the target object surface reflectance. 
Any variation resulting from the poisson nature of the photoreceptor response or due the image rendering software is negligible, and can be ignored.

(DO WE WANT TO GET AN ESTIMATE OF THE CLOUD SIZE FROM THE ALLOWED VARIATION IN THE SPECTRAL SHAPE?)

%% Figure 9
\begin{figure}
\centering
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_a.pdf}
        \caption{RF response Case 1}
        \label{fig:case1RFResponse}
    \end{subfigure}    
%    \begin{subfigure}[b]{0.33 \textwidth}   
 %       \includegraphics[width=\textwidth]{../Figures/resultsFigure/renderingNoise.pdf}
%        \caption{Rendering noise}
%        \label{fig:renderingNoise}
%    \end{subfigure}
            \begin{subfigure}[b]{0.297 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_b.pdf}
        \caption{Luminance Estimates}
        \label{fig:case9Results}
    \end{subfigure}    
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure9/Figure9_c.pdf}
        \caption{Images with fixed relative spectrum}
        \label{fig:case0RFResponse}
    \end{subfigure}
    \caption{{\bf Case 1: RF response and luminance estimates:} (a) Photoreceptor response projected along the first two AMA receptive fields for images of case 1. We show the response for the images in the training set. Each response cloud represents the response of images corresponding to one luminance level. The response clouds are approximated by a multinormal Gaussian whose mean and variance are approximated by the ellipses shown in the figure. (b) Mean luminance estimates and standard deviations for the images in the test set obtained using the naive method, linear regression and AMA. The estimates fall along the diagonal unit slope line for both linear regression and AMA. The standard deviations are too small to be visible. (c) Photoreceptor response for images with fixed relative shape of the target object reflectance spectrum projected along the first two case 1 AMA receptive fields. The variation coming from the poisson nature of the photoreceptor isomerization is small compared to the variation due to shape of target object reflectance spectrum.}
\label{fig:sourcesOfNoise}
\end{figure}

\subsection{Case 1a: This will be the case of background variation with fixed luminance}

\subsection{Case 2: Luminance can be estimated from contrast between target and background}
The AMA receptive fields learnt using the photoreceptor response does a very poor job in predicting the target object luminance (Fig.~\ref{fig:isomerizationFails}) for images of case 2. 
This is because, since the illuminant spectra varies in this case, the photoreceptor response is not directly proportional to the target object surface anymore. 
But, since the background surfaces are fixed in this case, the ratio between the response of  photoreceptors corresponding to the target object and the background remains proportional to the target object luminance. 
This is because a change in illuminant spectra produces the same effect in both the target object and the background object. 
Thus, the AMA receptive fields learnt on the contrast normalized photoreceptor response (see Methods) can be used to predicts the target object luminance (Fig.~\ref{fig:constrastWorks}). 
Fig.~\ref{fig:case10Results} shows the estimated target object luminance v/s the assigned luminance for the images in case 2 obtained using the contrast normalized representation. 
AMA receptive fields provide excellent predictions of the target object luminance. 
The predictions are much better compared to that of a linear model that only considers the (contrast normalized) response of the photoreceptors corresponding to the center pixel of the target object.

%% Figure Isomerization Fails
\begin{figure}
\centering
    \begin{subfigure}[b]{0.3	 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure10/Figure10_a.pdf}
        \caption{RFs learnt using photoreceptor response}
        \label{fig:isomerizationFails}
    \end{subfigure}
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure10/Figure10_b.pdf}
        \caption{RFs learnt using contrast signal}
        \label{fig:constrastWorks}
    \end{subfigure}
            \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure10/Figure10_c.pdf}
        \caption{Luminance estimates}
        \label{fig:case10Results}
    \end{subfigure}
    \caption{{\bf Case 2: Luminance can be estimated using the contrast signal:} (a) Photoreceptor response projected on AMA receptive fields calculated using photoreceptor response. (b) The contrast normalized photoreceptor response projected along the first two AMA receptive fields. The receptive fields were calculated using the contrast normalized stimuli. The response at each luminance level separates much better in this representation. (c) Estimated v/s assigned target object luminance for the image in case 2. Solid lines show the mean estimate, the filled region in light color  shows one standard deviation. AMA estimates are much better compared to the null models.}
\label{fig:importanceOfConstrast}
\end{figure}

While, the projection along the first two receptive fields shows an overlap between the stimuli at higher luminance values (Fig.~\ref{fig:constrastWorks}), the stimuli are better separated if the projection is considered within a subspace that has more receptive fields. 
Fig.~\ref{fig:RMSEvsNFilters} shows the relative error (Methods) of estimates as a function of the of the number of receptive fields considered for luminance estimation for the cases considered in this work. 
As expected, relative error reduces with increasing number of receptive fields.

%% Figure 11
\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{../Figures/Figure11/Figure11.pdf}
\caption{{\bf Performance with number of RFs used for estimation:} Most of the information is captured by the first two filters. We have used the first 8 RFs for estimating the luminance.}
\label{fig:RMSEvsNFilters}
\end{figure}
% GET RID OF AMA FROM THE CAPTION

\subsection{Case 3}
% Figure 12
\begin{figure}
\centering
    \begin{subfigure}[b]{0.3 \textwidth}   
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_a.pdf}
        \caption{AMA RF Response}
        \label{fig:case12FiltersResponse}
    \end{subfigure}    
        \begin{subfigure}[b]{0.3 \textwidth}
        \includegraphics[width=\textwidth]{../Figures/Figure12/Figure12_b.pdf}
        \caption{Luminance Estimates}
        \label{fig:case12Results}
    \end{subfigure}
    \caption{{\bf Case 3:} (a) Contrast normalized photoreceptor response projected along the first two AMA receptive fields. (b) Target object luminance estimates v/s assigned luminance for images in case 3. The AMA estimates are unbiased and have smaller relative error compared to linear model.}
    \label{fig:case12AllResults}
\end{figure}

Fig.~\ref{fig:case12AllResults} shows the AMA receptive field response and the luminance estimates for the images in case 3. 
We have used the first 8 receptive fields to estimate the luminance in this case (see Fig.\ref{fig:RMSEvsNFilters}). 
The AMA receptive fields do an excellent job of predicting the target object luminance. The estimates are unbiased with the mean estimates lying close the diagonal unit slope line (Fig.~\ref{fig:case12Results}). 
The relative error of estimation is ~17\%, which low compared to that of the naive model (~42\%) or the linear model on center pixel (~21\%).

@@

\subsection{The RFs of complex case generalize on simpler cases}
Fig.~\ref{fig:barGraphs} shows the relative error for the estimates of the target object luminance obtained using the linear model and AMA. We show the relative error of the estimates obtained using the AMA receptive fields from all cases. The receptive fields from a more complex case perform at par on the stimuli from a simpler case. The optimal AMA estimates are better than that of the linear model.

% Figure 13
\begin{figure}
\centering
\begin{subfigure}{0.3 \textwidth}
	\includegraphics[width=\textwidth]{../Figures/Figure13/Figure13_a.pdf}
	\caption{Case 1}
	\label{fig:case1Bar}
    \end{subfigure}
    ~ 
    \begin{subfigure}{0.3 \textwidth}   
	\includegraphics[width=\textwidth]{../Figures/Figure13/Figure13_b.pdf}
	\caption{Case 2}
	\label{fig:case2Bar}
    \end{subfigure}
    ~ 
        \begin{subfigure}{0.3 \textwidth}
	\includegraphics[width=\textwidth]{../Figures/Figure13/Figure13_c.pdf}
	\caption{Case 3}
	\label{fig:case3Bar}
    \end{subfigure}
\caption{{\bf Error comparison across models:} Comparison of optimal AMA receptive fields from one case on the images of other cases. (a) Relative error of the estimates obtained using the linear model and AMA. The AMA receptive fields of case 1, 2 and 3 are used to estimate the target object luminance of the test images from case 1. The AMA receptive fields from all three cases perform well in this case. (b) Relative error for images of case 2. As expected, the error of the linear model on the center pixel is higher than AMA. AMA receptive fields from case 1 do worse than the receptive fields from case 2 and 3. (c) Relative error for images of case 3. The best predictions are from the receptive learnt using the images of case 3.}
 \label{fig:barGraphs}
\end{figure}

@@
